A fast and Low-Overhead Technique to Secure Programs Against Integer Overflows

### 2012-09-06 doc/Manuscript/CGO/paper.tex
Experimental Section

\noindent
\textbf{Time and Memory Complexity.}
Figure~\ref{fig:TimeCorr} provides a visual comparison between the time to
run our range analysis and the size of the input programs.
We show data for the 100 largest benchmarks in our test suite, in number
of variable nodes in the constraint graph.
We perform function inlining before running our analysis.
Each point in the X line corresponds to a benchmark.
We analyze the smallest benchmark in this set, \texttt{Prolangs-C/deriv2}, which
has 1,131 variable nodes in the constraint graph, in 20ms.
We take 15.91 sec to analyze our largest benchmark, \texttt{403.gcc}, which,
after function inlining, has 1,266,273 assembly instructions, and a
constraint graph with 679,652 variable nodes.
For this data set, the coefficient of determination $(R^2)$ is 0.967, which
provides very strong evidence about the linear asymptotic complexity of our
implementation.

@@

\noindent
\textbf{Precision.}
Our implementation of range analysis is remarkably precise, considering its
runtime.
Lakhdar {\em et al.}'s relational analysis~\cite{Lakhdar11}, for instance, takes
about 25 minutes to go over a program with almost 900 basic blocks.
We analyze programs of similar size in less than one second.
We do not claim our approach is as precise as such algorithms, even though we
are able to find exact bounds to 4/5 of the examples presented
in~\cite{Lakhdar11}.
On the contrary, we present a compromise between precision and speed
that scales to very large programs.
Nevertheless, our results are far from being trivial.
We have implemented a dynamic profiler that measures, for each variable,
its upper and lower limits, given an execution of the target program.
Figure~\ref{fig:precision} compares our results with those measured
dynamically for the Stanford benchmark, which is publicly
available in the LLVM test suite.
We chose Stanford because these benchmarks do not read data from external
files; hence, imprecisions are due to library functions that we cannot
analyze.

@@

%How we have done the dynamic tests  
We have executed the instrumented programs of the integer benchmarks of SPEC
2006 CPU to measure the overhead produced by our instrumentation in 
real world programs.
To do this, we have developed a little test framework that compiles, runs and
collects the statistics about the execution of all the programs of SPEC 2006.
The execution of the programs was made using the arguments and inputs of the "test"
mode, which is faster to execute than the "ref" inputs.

%Analysis about the Table 1 of the spreadsheet (Unpruned Instrumentation)
Figure~\ref{fig:unpruned_analysis} shows the number of instructions that has
received instrumentation, without avoiding inserting the instrumentation by
using any static analysis. We can see that the number of these instructions is low 
(3,96 in the average). 
Figure~\ref{fig:prunning_analysis} shows that it is possible to use static 
analysis to avoid inserting the instrumentation in some instructions.

@@

\noindent
\paragraph{Dynamic Detection of Integer Overflows: }
Brumley {\em et al.}~\cite{Brumley07} have developed a tool, RICH, to
dynamically detect integer overflows.
Their goal is to secure C programs against integer overflow based
vulnerabilities.
The author's approach consists in instrumenting every integer operation that
might cause an overflow, underflow, or data loss.
The main result of Brumley {\em et al.} is the verification that instrumenting
programs to guard them against integer overflows does not compromise their
performance significantly: the average slowdown across four large applications
is 5\%.
RICH, Brumley {\em et al}'s tool, uses specific features of the x86 architecture
to reduce the instrumentation overhead.
Chinchani {\em et al.}~\cite{Chinchani04} follow a similar approach.
In this work, the authors describe each arithmetic operation formally, and then
use characteristics of the computer architecture to detect overflows at
runtime.
% TODO bibtex: cite Archerr: Runtime environment driven program safety.
Contrary to these previous work, we instrument programs at LLVM's intermediate
representation level, which is machine independent.
Nevertheless, the performance of the programs that we instrument is on par with
Brumley's, even without the support of the static range analysis.

@@

\paragraph{Static Detection of Integer Overflows: }
Zhang {\em et al.}~\cite{Zhang09} have used static analysis to sanitize
programs against integer overflow based vulnerabilities.
They instrument integer operations in paths from a source to a sink.
In Zhang {\em et al.}'s context, sources are functions that read values from
users, and sinks are memory allocation operations.
Thus, contrary to our work, Zhang {\em et al.}'s only need to instrument
about 10\% of the integer operations in the program.
However, they do not use any form of range analysis to limit the number of
checks inserted in the transformed code.
Wang {\em et a.}~\cite{Wang09} have implemented a tool, IntScope, that combines symbolic execution and taint analysis to detect integer overflow vulnerabilities. The authors have been able to use this tool to successfully identify many vulnerabilities in industrial quality software. Our work, and Wang {\em et a.}'s work are essentially different: they use symbolic execution, whereas we rely on
range analysis.
Furthermore, contrary to them, we do not try to detected statically the
possibility of an integer overflow happening.
Contrary to us, they do not transform the program to prevent or detect such event dynamically.
Still in the field of symbolic execution, Molnar
{\em et al.}~\cite{Molnar09} have implemented a tool, SmartFuzz, that analyzes
Linux x86 binaries to find integer overflow bugs.
They prove the existence of bugs by generating test cases for them.

\paragraph{Range Analysis:}
One of the goals of this paper is to describe a state-of-the-art range analysis
algorithm.
Many parts of this algorithm are well-known, others are novel.
The widening and narrowing operators from Section~\ref{sub:micro} were
introduced by Cousot and Cousot's seminal paper~\cite{Cousot77}.
The widening operator that we use in our implementation is well-known among
compiler writers~\cite[p.228]{Nielson99}.
The interval lattice from Section~\ref{sub:lattice} has been used in a
plethora of old and recent range analysis
solvers~\cite{Gough94,Mahlke01,Oh12,Patterson95,Stephenson00,Su05}.
The use of futures to handle comparisons between variables is a contribution
of this paper.
It is well-known that the use of strongly connected components speeds up
constraint resolution systems~\cite[Sec 6.3]{Nielson99}.
However, publicly available implementations of range analysis such as those in
FLEX~\footnote{The MIT's FLEX/Harpoon compiler
provides an implementation of Stephenson's algorithm~\cite{Stephenson00}, and
is available at \texttt{http://flex.cscott. net/Harpoon/}.},
gcc~\footnote{Gcc's VRP pass (at \texttt{http://gcc.gnu.org/svn/gcc/trunk/gcc/
tree-vrp.c}) implements a variant of Patterson's
algorithm~\cite{Patterson95}.} or Mozilla's IonMonkey~\footnote{\texttt{hg.mozilla.org/projects/ionmonkey/file/629d1e6251b9/js/
src/ion/RangeAnalysis.cpp}. This code is implemented after Gough and
Klaeren~\cite{Gough94}} do not resort to this technique.
Furthermore, they only deal with comparisons between variables and constants.

\noindent
\paragraph{Sparse Dataflow Analysis: }
In this paper we provide a {\em sparse} implementation of range analysis.
Sparsity, in our context, means that we associate points in the lattice of
interest -- intervals in our case -- directly to variables.
Dense analyses map such information to pairs formed by variables and program
points.
The compiler related literature contains many descriptions of sparse
data-flow analyses.
Some analyses obtain sparsity by using specific program
representations, like we did.
Others rely on data-structures.
In terms of data-structures, the first, and best known method proposed to
support sparse data-flow analyses is Choi {\em et al.}'s {\em Sparse Evaluation
Graph} (SEG)~\cite{Choi91}.

One of the program representation that we discuss in
Section~\ref{sub:splitting} -- the {\em
Extended Static Single Assignment} (e-SSA) form -- has been used by Bodik
{\em et al.}~\cite{Bodik00} and Gough {\em et al.}~\cite{Gough94}.
The other representation, which we call u-SSA, is very similar to
the \emph{Static Single Use} form (SSU).
There exists many variants of SSU~\cite{Plevyak96,George03,Lo98}.
The main difference is that our dataflow analysis propagates information
forwardly, whereas the SSU form has been conceive for dataflow analyses
that propagate information backwardly.
Thus, it requires special nodes at branches to merge information from
different program paths.
The SSU form programs are a subset of the programs in the so called
{\em Static Single Information} form, a program representation proposed by
Scott Ananian~\cite{Ananian99}.
SSI makes sparse the dataflow analyses that propagate information both
forwardly and backwardly.

@@

>>>>

\subsection{Static Range Analysis}
\label{sub:range}

\noindent
\textbf{Time and Memory Complexity: }
Figure~\ref{fig:TimeCorr} provides a visual comparison between the time to
run our range analysis and the size of the input programs.
We show data for the 100 largest benchmarks in our test suite, in number
of variable nodes in the constraint graph.
We perform function inlining before running our analysis.
Each point in the X line corresponds to a benchmark.
We analyze the smallest benchmark in this set, \texttt{Prolangs-C/deriv2}, which
has 1,131 variable nodes in the constraint graph, in 20ms.
We take 15.91 sec to analyze our largest benchmark, \texttt{403.gcc}, which,
after function inlining, has 1,266,273 assembly instructions, and a
constraint graph with 679,652 variable nodes.
For this data set, the coefficient of determination $(R^2)$ is 0.967, which
provides very strong evidence about the linear asymptotic complexity of our
implementation.

@@

\noindent
\textbf{Precision: }
Our implementation of range analysis offers a good tradeoff between
precision and runtime.
Lakhdar {\em et al.}'s relational analysis~\cite{Lakhdar11}, for instance,
takes about 25 minutes to go over a program with almost 900 basic blocks.
We analyze programs of similar size in less than one second.
We do not claim our approach is as precise as such algorithms, even though we
are able to find exact bounds to 4/5 of the examples presented
in~\cite{Lakhdar11}.
On the contrary, we present a compromise between precision and speed
that scales to very large programs.
Nevertheless, our results are far from being trivial.
We have implemented a dynamic profiler that measures, for each variable,
its upper and lower limits, given an execution of the target program.
Figure~\ref{fig:precision} compares our results with those measured
dynamically for the Stanford benchmark, which is publicly
available in the LLVM test suite.
We chose Stanford because these benchmarks do not read data from external
files; hence, imprecisions are due to library functions that we cannot
analyze.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Static Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%How we have done the dynamic tests
\subsection{The Instrumentation Library}
\label{sub:inst}

We have executed the instrumented programs of the integer benchmarks of SPEC
2006 CPU to probe the overhead imposed by our instrumentation.
All the programs have been executed with its ``test" input set.
We have not been able to run the binary that LLVM produces for SPEC's
\texttt{gcc} in our environment, even without any of our passes.
Figure~\ref{fig:unprunedTable} shows the percentage of instructions that
we instrument, without the intervention of the range analysis.
The number of instrumented instructions is relatively low, compared to the
total number of instructions, because we only instrument six different
LLVM bitcodes, in a set of 57 opcodes, not counting intrinsics.
Figure~\ref{fig:unprunedTable} also shows how many instructions have caused
overflows.
On the average, 4.90\% of the instrumented sites have caused integer overflows.

%Analysis about the amount of instructions
\begin{figure}[htb]
\begin{center}
\begin{small}
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|}
\hline
Benchmark & \#I & \#II & \#II/\#I & \#O \\ \hline
470.lbm & 13,724 & 1,142 & 8.32\% & 0 \\ \hline
433.milc & 44,236 & 1,602 & 3.62\% & 11 \\ \hline
444.namd & 100,276 & 3,234 & 3.23\% & 12 \\ \hline
447.dealII & 1,381,408 & 36,157 & 2.62\% & 50 \\ \hline
450.soplex & 136,367 & 3,158 & 2.32\% & 13 \\ \hline
464.h264ref & 271,627 & 13,846 & 5.10\% & 167 \\ \hline
473.astar & 19,243 & 857 & 4.45\% & 0 \\ \hline
458.sjeng & 54,051 & 2,504 & 4.63\% & 68 \\ \hline
429.mcf & 4,725 & 165 & 3,49\% & 8 \\ \hline
471.omnetpp & 203,201 & 1,972 & 0.97\% & 2 \\ \hline
403.gcc & 1,419,456 & 18,669 & 1.32\% & N/A \\ \hline
445.gobmk & 308,475 & 14,129 & 4.58\% & 4 \\ \hline
462.libquantum & 16,297 & 928 & 5.69\% & 7 \\ \hline
401.bzip2 & 38,831 & 2,158 & 5.56\% & N/A \\ \hline
456.hmmer & 114,136 & 4,001 & 3.51\% & 0 \\ \hline
\multicolumn{1}{|r|}{Total (Average)} & 275.070 & 6.968 & 3,96\% &  \\ \hline
\end{tabular*}
\end{small}
\end{center}
\caption{\label{fig:unprunedTable}
Instrumentation library without support of static range analysis.
\#I: number of LLVM bitcode instructions in the original program.
\#II: number of instructions that have been instrumented.
\#O: number of actual overflow tests detected dynamically.}
\end{figure}

Figure~\ref{fig:prunningTable} shows how many checks our range analysis
can avoid.
In some benchmarks we obtain impressive results:
For instance, the range analysis avoids 1,138 out of 1,142 checks in
\texttt{470.lbm}.
In other benchmarks, such as in \texttt{429.mcf}, we have been able to
avoid only 1, out of 165 tests.
In general we fare better in programs that bound input sizes via conditional
tests, as \texttt{lbm} does.
Using u-SSA, instead of e-SSA adds a negligible improvement onto our results.
We speculate that this improvement is small because variables tend to be used
a small number of times.
Benoit {\em et al.}~\cite{Benoit08} have demonstrated that the vast majority
of all the program variables are used less than five times in the program
code.
The u-SSA form only helps to avoid checks upon variables that are used more
than once.

%Prunning analysis
\begin{figure}[t!]
\begin{center}
\begin{small}
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|r|}
\hline
Benchmark & \#II & \#E & \#I/\#E & \#U &
\#I/\#U \\ \hline
470.lbm & 1.142 & 4 & 99,65\% & 4 & 99,65\% \\ \hline
433.milc & 1.573 & 1.070 & 31,98\% & 1.065 & 32,29\% \\ \hline
444.namd & 2.965 & 2.900 & 2,19\% & 2.900 & 2,19\% \\ \hline
447.dealII & 37.234 & 29.870 & 19,78\% & 28.779 & 22,71\% \\ \hline
450.soplex & 3.268 & 2.927 & 10,43\% & 2.897 & 11,35\% \\ \hline
464.h264ref & 13.548 & 11.342 & 16,59\% & 11.301 & 16,28\% \\ \hline
473.astar & 903 & 808 & 10,52\% & 806 & 10,74\% \\ \hline
458.sjeng & 2.483 & 2.354 & 5,20\% & 2.190 & 11,80\% \\ \hline
429.mcf & 165 & 164 & 0,61\% & 164 & 0,61\% \\ \hline
471.omnetpp & 1.533 & 1.313 & 14,35\% & 1.313 & 14,35\% \\ \hline
403.gcc & 16.330 & 15.282 & 6,42\% & 15.110 & 7,47\% \\ \hline
445.gobmk & 13.947 & 12.563 & 9,92\% & 12.478 & 10,53\% \\ \hline
462.libquantum & 943 & 820 & 13,04\% & 817 & 13,36\% \\ \hline
401.bzip2 & 2.350 & 1.966 & 16,34\% & 1.966 & 16,34\% \\ \hline
456.hmmer & 3.963 & 3.346 & 15,57\% & 3.304 & 16,63\% \\ \hline
Total (Average) & 6.823 & 5.779 & 18,17\% & 5.676 & 19,09\% \\ \hline
\end{tabular*}
\end{small}
\end{center}
\caption{\label{fig:prunningTable}
Instrumentation library with support of static range analysis.
\#II: number of instructions that have been instrumented without range
analysis.
\#E: number of instructions instrumented in the e-SSA form program.
\#U: number of instructions instrumented in the u-SSA form program.}
\end{figure}

Figure~\ref{fig:classificationTable} shows how our range analysis
classifies instructions.
Out of all the 102,790 instructions that we have instrumented throughout the
entire SPEC suite, 3.92\% are suspicious, 17.19\% are safe, and
78.89\% are uncertain.
We had, at first, speculated that overflows would be more common among
suspicious instructions, as their bounds, inferred statically, go beyond the
limits of their declared types.
However, our experiments did not let us confirm this hypothesis.

%Classification analysis
\begin{figure}[t!]
\begin{center}
\begin{scriptsize}
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
\hline
Bench & \#Sf & \#S & \#U & \#SO & \#SO/\#O & \#UO & \#UO/\#O
\\ \hline
lbm     & 1138 & 0    & 4 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
milc    & 536  & 17   & 1048  & 0 & 0,00\% & 11 & 1,05\% \\ \hline
namd    & 334  & 480  & 2420  & 0 & 0,00\% & 12 & 0,50\% \\ \hline
dealII  & 6188 & 39   & 28740  & 0 & 0,00\% & 50 & 0,17\% \\ \hline
soplex  & 229  & 16   & 2881 & 0 & 0,00\% & 13 & 0,45\% \\ \hline
h264ref & 2539 & 1195 & 10147 & 7 & 0,59\% & 167 & 1,65\% \\ \hline
astar   & 48   & 11   & 795 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
sjeng   & 150  & 213  & 1977 &0 & 0,00\% & 68 & 3,44\% \\ \hline
mcf     & 1    & 0    & 164 & 0 & 0,00\% & 8 & 4,88\% \\ \hline
omnetpp & 659  & 25   & 1288 & 1 & 4,00\% & 2 & 0,16\% \\ \hline
gcc     & 3365 & 1045 & 14065 & N/A & N/A & N/A & N/A \\ \hline
gobmk   & 1509 & 742  & 11736 & 0 & 0,00\% & 4 & 0,03\% \\ \hline
libqtum & 104  & 12   & 805 & 0 & 0,00\% & 7 & 0,87\% \\ \hline
bzip2   & 192  & 40   & 1926 & N/A & N/A & N/A & N/A \\ \hline
hmmer   & 663  & 222  & 3082 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
\end{tabular}
\end{scriptsize}
\end{center}
\caption{\label{fig:classificationTable}
How the range analysis classified arithmetic instructions.
\#Sf: safe.
\#S: suspicious.
\#U: uncertain.
\#SO: number of suspicious instructions that overflowed.
\#UO: number of uncertain instructions that overflowed.
}
\end{figure}

Figure~\ref{fig:percentage_elimination} shows, for the entire LLVM test
suite, the percentage of overflow checks that our range analysis, with the
e-SSA intermediate representation, could avoid.
Each bar refers to a specific benchmark in the test suite.
We only consider applications that had at least one instrumented
instruction; the total number of benchmarks that meet this requirement is
333.
On the average, our range analysis avoid 24.93\% of the overflow checks.
Considering the benchmarks in SPEC2006 only, this number is 20.57\%.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\columnwidth]{images/percentage_elimination}
\end{center}
\caption{\label{fig:percentage_elimination}
Percentage of overflow checks that our range analysis has been able to remove.
Each bar is a benchmark in the LLVM test suite.
Benchmarks have been ordered by the effectiveness of the range analysis.
On average, we have eliminated 24.93\% of the checks (geomean).}
\end{figure}

Figure~\ref{fig:spec_runtime} shows the impact of our instrumentation in the
runtime of the SPEC benchmarks.
We run each benchmark 20 times.
The largest slowdown that we have observed, 11.83\%, happened in
\texttt{h264ref}, the benchmark that presented the largest number of distinct
sites where overflows happened dynamically. 
On the average, the instrumented programs are 3.24\% slower than the
original benchmarks.
If we use the range analysis to eliminate overflow checks, this slowdown falls
to 1.73\%.
The range analysis, in this case, reduces the instrumentation overhead by
53.40\%.
This improvement is larger than the percentage of overflow checks that we
avoid, e.g,. 20.57\%.
We believe that this difference is due to the fact that we are able to
eliminate checks on induction variables, as our range analysis can rely on
the loop boundaries to achieve this end.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=\columnwidth]{images/spec_runtime}
\end{center}
\caption{\label{fig:spec_runtime}
Comparison between the execution times with and without prunning, 
normalized by the original program's execution time.}
\end{figure}

@@

\noindent
\textbf{Dynamic Detection of Integer Overflows: }
Brumley {\em et al.}~\cite{Brumley07} have developed a tool, RICH, to
dynamically detect integer overflows.
Their goal is to secure C programs against integer overflow based
vulnerabilities.
The author's approach consists in instrumenting every integer operation that
might cause an overflow, underflow, or data loss.
The main result of Brumley {\em et al.} is the verification that instrumenting
programs to guard them against integer overflows does not compromise their
performance significantly: the average slowdown across four large applications
is 5\%.
RICH, Brumley {\em et al}'s tool, uses specific features of the x86 architecture
to reduce the instrumentation overhead.
Chinchani {\em et al.}~\cite{Chinchani04} follow a similar approach.
In this work, the authors describe each arithmetic operation formally, and then
use characteristics of the computer architecture to detect overflows at
runtime.
% TODO bibtex: cite Archerr: Runtime environment driven program safety.
Contrary to these previous work, we instrument programs at LLVM's intermediate
representation level, which is machine independent.
Nevertheless, the performance of the programs that we instrument is on par with
Brumley's, even without the support of the static range analysis.

@@

\noindent
\textbf{Static Detection of Integer Overflows: }
Zhang {\em et al.}~\cite{Zhang09} have used static analysis to sanitize
programs against integer overflow based vulnerabilities.
They instrument integer operations in paths from a source to a sink.
In Zhang {\em et al.}'s context, sources are functions that read values from
users, and sinks are memory allocation operations.
Thus, contrary to our work, Zhang {\em et al.}'s only need to instrument
about 10\% of the integer operations in the program.
However, they do not use any form of range analysis to limit the number of
checks inserted in the transformed code.
Wang {\em et a.}~\cite{Wang09} have implemented a tool, IntScope, that combines symbolic execution and taint analysis to detect integer overflow vulnerabilities. The authors have been able to use this tool to successfully identify many vulnerabilities in industrial quality software. Our work, and Wang {\em et a.}'s work are essentially different: they use symbolic execution, whereas we rely on
range analysis.
Furthermore, contrary to them, we do not try to detected statically the
possibility of an integer overflow happening.
Contrary to us, they do not transform the program to prevent or detect such event dynamically.
Still in the field of symbolic execution, Molnar
{\em et al.}~\cite{Molnar09} have implemented a tool, SmartFuzz, that analyzes
Linux x86 binaries to find integer overflow bugs.
They prove the existence of bugs by generating test cases for them.

\noindent
\textbf{Range Analysis and sparse representations:}
One of the goals of this paper is to describe a state-of-the-art range analysis
algorithm.
Many parts of this algorithm are well-known, others are novel.
The widening and narrowing operators from Section~\ref{sub:micro} were
introduced by Cousot and Cousot's seminal paper~\cite{Cousot77}.
The widening operator that we use in our implementation is well-known among
compiler writers~\cite[p.228]{Nielson99}.
The interval lattice from Section~\ref{sub:lattice} has been used in a
plethora of old and recent range analysis
solvers~\cite{Gough94,Mahlke01,Oh12,Patterson95,Stephenson00,Su05}.
The use of futures to handle comparisons between variables is a contribution
of this paper.
It is well-known that the use of strongly connected components speeds up
constraint resolution systems~\cite[Sec 6.3]{Nielson99}.
However, publicly available implementations of range analysis such as those in
FLEX~\footnote{The MIT's FLEX/Harpoon compiler
provides an implementation of Stephenson's algorithm~\cite{Stephenson00}, and
is available at \texttt{http://flex.cscott. net/Harpoon/}.},
gcc~\footnote{Gcc's VRP pass (at \texttt{http://gcc.gnu.org/svn/gcc/trunk/gcc/
tree-vrp.c}) implements a variant of Patterson's
algorithm~\cite{Patterson95}.} or Mozilla's IonMonkey~\footnote{\texttt{hg.mozilla.org/projects/ionmonkey/file/629d1e6251b9/js/
src/ion/RangeAnalysis.cpp}. This code is implemented after Gough and
Klaeren~\cite{Gough94}} do not resort to this technique.
Furthermore, they only deal with comparisons between variables and constants.
The only work that we are aware of that has used range analysis to avoid
overflow checks is Sol {\em et al}'s~\cite{Sol11}; however, their algorithm
is much more limited than ours.
Contrary to us, Sol {\em et al.} eliminate overflow checks in
straight-line code traces produced by a just-in-time compiler, and they use the
runtime values of variables to improve their precision.

In this paper we provide a {\em sparse} implementation of range analysis.
Sparsity, in our context, means that we associate points in the lattice of
interest -- intervals in our case -- directly to variables.
Dense analyses map such information to pairs formed by variables and program
points.
Some previous analyses have achieved sparsity by using specific program
representations, like we did.
One of the program representation that we discuss in
Section~\ref{sub:splitting} -- the {\em
Extended Static Single Assignment} (e-SSA) form -- has been used by Bodik
{\em et al.}~\cite{Bodik00} and Gough {\em et al.}~\cite{Gough94}.
The other representation, which we call u-SSA, is very similar to
the \emph{Static Single Use} form (SSU).
There exists many variants of SSU~\cite{Plevyak96,George03,Lo98}.
The main difference to our representation is that our dataflow analysis
propagates information forwardly, whereas the SSU form has been conceive for
dataflow analyses that propagate information backwardly.
Thus, it requires special nodes at branches to merge information from
different program paths.
The SSU form programs are a subset of the programs in the so called
{\em Static Single Information} form, a program representation proposed by
Scott Ananian~\cite{Ananian99}.
SSI makes sparse the dataflow analyses that propagate information both
forwardly and backwardly.

@@

This paper has presented a static range analysis algorithm that reduces
the overhead necessary to secure programs against integer overflows.
This algorithm is able to analyze inter-procedurally programs with
half-a-million lines of code, which produce almost one million constraints, in
ten seconds.
We proposed the notion of ``future bounds" as a way to handle comparisons
between variables, and relied on strong connected components to scale up to
large programs.
Although the overhead of guarding programs against integer overflows is small,
as previous work has demonstrated, we believe that our technique is still
important, as its gains are cumulative with improvements in hardware.
Our entire code is publicly available, as a series of LLVM patches.
The data that we show in this paper is also anonymously available, to ensure
the reproducibility of our experiments.
